---
title: "Visualization of inequality"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
    runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(reshape2)
library(plotly)

```

# HDR

## ##row {.tabset data-width="650"}

### Human development group

```{r}
file_path <- "D:/DV (RStudio)/project_dataset.csv"

df <- read.csv(file_path)
df_filtered <- df %>% filter(!is.na(Human.Development.Groups))
count_data <- df_filtered %>% count(Human.Development.Groups)

ggplot(count_data, aes(x = reorder(Human.Development.Groups, -n), y = n, fill = Human.Development.Groups)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Low" = "coral", "Medium" = "yellow", "High" = "lightblue", "Very High" = "pink")) +
  labs(title = "Distribution of Human Development Groups (Excluding NA)",
       x = "Human Development Groups",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Distribution of Human Development Groups

```{r}
data <- data.frame(
  Human.Development.Groups = c("Low", "Medium", "High", "Very High"),
  Count = c(49, 44, 4, 66)  # Replace with actual counts
)

# Create a pie chart
ggplot(data, aes(x = "", y = Count, fill = Human.Development.Groups)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Distribution of Human Development Groups")

```


### HDR
```{r}
selectInput(inputId = "type", label = "Human development group", 
            choices = c("Low", "Medium", "High","Very high"))


renderPlot({
  if (input$type == "Low") {
    low_hdi_df <- df[df$Human.Development.Groups == "Low", ]
    
    # Create a bar plot
    ggplot(data = low_hdi_df, aes(x = reorder(Country, HDI.Rank..2021.), y = HDI.Rank..2021.)) +
      geom_bar(stat = "identity", fill = "lightblue") +
      coord_flip() +
      labs(title = "Countries with Low Human Development",
           x = "HDI Rank (2021)",
           y="Country")
  } 
  
  
  else  if (input$type == "Medium") {
    low_hdi_df <- df[df$Human.Development.Groups == "Medium", ]
    
    # Create a bar plot
    ggplot(data = low_hdi_df, aes(x = reorder(Country, HDI.Rank..2021.), y = HDI.Rank..2021.)) +
      geom_bar(stat = "identity", fill = "pink") +
      coord_flip() +
      labs(title = "Countries with Medium Human Development",
           x = "HDI Rank (2021)",
           y="Country")
  } 
  
  
  
  else if (input$type == "High") {
    low_hdi_df <- df[df$Human.Development.Groups == "High", ]
    
    # Create a bar plot
    ggplot(data = low_hdi_df, aes(x = reorder(Country, HDI.Rank..2021.), y = HDI.Rank..2021.)) +
      geom_bar(stat = "identity", fill = "red") +
      coord_flip() +
      labs(title = "Countries with High Human Development",
           x = "HDI Rank (2021)",
           y="Country")
  } 
  else  if (input$type == "Very high") {
    low_hdi_df <- df[df$Human.Development.Groups == "Very High", ]
    
    # Create a bar plot
    ggplot(data = low_hdi_df, aes(x = reorder(Country, HDI.Rank..2021.), y = HDI.Rank..2021.)) +
      geom_bar(stat = "identity", fill = "green") +
      coord_flip() +
      labs(title = "Countries with Very High Human Development",
           x = "HDI Rank (2021)",
           y="Country")
  } 
})

```
# Map {data-icon="fa-map"}

## column {data-width="650"}

### Map

```{r}
data <- df 

plot_map <- plot_ly(
  data,
  z = ~HDI.Rank..2021., 
  locations = ~Country,
  type = "choropleth",
  locationmode = "country names"
) %>%
  colorbar(title = "HDI Rank 2021")  

plot_map %>%
  layout(
    geo = list(
      showframe = FALSE,
      showcoastlines = TRUE,
      projection = list(type = "mercator")
    ),
    title = "Choropleth Map of HDI Rank 2021 by Country"
  )
```

# Percentage {data-icon="fa-percentage"}

### Colored Scatter Plot of Percentage of Inequality by Year.

```{r}
sorted_df <- df %>% arrange(HDI.Rank..2021.)
sum_inequality <- colSums(sorted_df[, 6:17], na.rm = TRUE)
total_inequality_count <- sum(sum_inequality, na.rm = TRUE)
percentage_inequality <- (sum_inequality / total_inequality_count) * 100
inequality_summary <- data.frame(
  Year = names(sum_inequality),
  Total_Count = sum_inequality,
  Percentage = percentage_inequality
)
plot_ly(inequality_summary, x = ~Year, y = ~Percentage, color = ~Year, type = 'scatter', mode = 'markers') %>%
  layout(title = "Percentage of Inequality by Year",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Percentage"),
         showlegend = FALSE)



```

# Covid

```{r}
data <- df 
data$Change <- data$`Inequality.in.Education..2021.` - data$`Inequality.in.Education..2019.`

scatter_plot <- plot_ly(data, x = ~`Inequality.in.Education..2019.`, y = ~Change, text = ~Country,
                        type = 'scatter', mode = 'markers', marker = list(size = 10))

layout(scatter_plot) %>%
  layout(
    title = "Change in Inequality pre and post Covid (2019 to 2021)",
    xaxis = list(title = "Inequality in Education 2019"),
    yaxis = list(title = "Change in Inequality (2019 to 2021)")
  )

slider <- list(
  currentvalue = list(prefix = "Year: "),
  steps = list(
    list(
      args = list("x", list(data$`Inequality.in.Education..2019`), "y", list(data$Change)),
      method = "animate",
      label = "2019"
    ),
    list(
      args = list("x", list(data$`Inequality.in.Education..2021`), "y", list(data$Change)),
      method = "animate",
      label = "2021"
    )
  )
)

scatter_plot <- scatter_plot %>%
  layout(updatemenu = list(type = "buttons", showactive = FALSE, buttons = list(slider)))

scatter_plot

```

# Difference {data-icon="fa-clock"}

## row {.tabset .tabset-fade}

### (2020 vs. 2021)

```{r}
combined_data <- sorted_df %>%
  filter(!is.na(Inequality.in.Education..2020.), !is.na(Inequality.in.Education..2021.))

ggplot(combined_data, aes(x = "2020 vs. 2021", y = Inequality.in.Education..2021. - Inequality.in.Education..2020.)) +
  geom_boxplot() +
  labs(title = "Comparison of Inequality in Education (2020 vs. 2021)") +
  xlab("Comparison") +
  ylab("Difference in Inequality (2021 - 2020)") +
  theme_minimal()


```

### (2010 vs. 2021)

```{r}
combined_data <- sorted_df %>%
  filter(!is.na(Inequality.in.Education..2010.), !is.na(Inequality.in.Education..2021.))

combined_data$Difference <- combined_data$Inequality.in.Education..2021. - combined_data$Inequality.in.Education..2010.

top_15 <- head(combined_data[order(combined_data$Difference, decreasing = TRUE), ], 15)
bottom_15 <- head(combined_data[order(combined_data$Difference, decreasing = FALSE), ], 15)

top_and_bottom_15 <- rbind(top_15, bottom_15)

plot <- ggplot(top_and_bottom_15, aes(x = reorder(Country, -Difference), y = Difference, fill = ifelse(Difference >= 0, "Top", "Bottom"))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Top" = "skyblue", "Bottom" = "coral")) +
  labs(title = "Differences in Inequality in Education (2010 vs. 2021) for Top and
       Bottom 15 Countries") +
  xlab("Country") +
  ylab("Difference in Inequality") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot + guides(fill = guide_legend(title = NULL))



```

### (2015 vs. 2021)

```{r}
combined_data <- sorted_df %>%
  filter(!is.na(Inequality.in.Education..2015.), !is.na(Inequality.in.Education..2021.))

combined_data$Difference <- combined_data$Inequality.in.Education..2015. - combined_data$Inequality.in.Education..2021.

top_15 <- head(combined_data[order(combined_data$Difference, decreasing = TRUE), ], 15)
bottom_15 <- head(combined_data[order(combined_data$Difference, decreasing = FALSE), ], 15)

top_and_bottom_15 <- rbind(top_15, bottom_15)

plot <- ggplot(top_and_bottom_15, aes(x = reorder(Country, -Difference), y = Difference, fill = ifelse(Difference >= 0, "Top", "Bottom"))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Top" = "skyblue", "Bottom" = "coral")) +
  labs(title = "Differences in Inequality in Education (2010 vs. 2015) for Top and
       Bottom 15 Countries") +
  xlab("Country") +
  ylab("Difference in Inequality") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot + guides(fill = guide_legend(title = NULL))




```

### (2010 vs. 2015)

```{r}
combined_data <- sorted_df %>%
  filter(!is.na(Inequality.in.Education..2010.), !is.na(Inequality.in.Education..2015.))

combined_data$Difference <- combined_data$Inequality.in.Education..2010. - combined_data$Inequality.in.Education..2015.

top_15 <- head(combined_data[order(combined_data$Difference, decreasing = TRUE), ], 15)
bottom_15 <- head(combined_data[order(combined_data$Difference, decreasing = FALSE), ], 15)

top_and_bottom_15 <- rbind(top_15, bottom_15)

plot <- ggplot(top_and_bottom_15, aes(x = reorder(Country, -Difference), y = Difference, fill = ifelse(Difference >= 0, "Top", "Bottom"))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Top" = "skyblue", "Bottom" = "coral")) +
  labs(title = "Differences in Inequality in Education (2010 vs. 2015) for Top and
       Bottom 15 Countries") +
  xlab("Country") +
  ylab("Difference in Inequality") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot + guides(fill = guide_legend(title = NULL))

```

# 2021 {data-icon="fa-signal"}

## Row {.tabset .tabset-fade}

### Top 20 countries

```{r}
top_20_2021 <-  head(sorted_df[!is.na(sorted_df$Inequality.in.Education..2021.),], 20)

ggplot(top_20_2021, aes(x = reorder(Country, Inequality.in.Education..2021.), y = Inequality.in.Education..2021.)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Inequality in Education by Country (2021)") +
  xlab("Country") +
  ylab("Inequality in Education (2021)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Bottom 20 countries

```{r}
bottom_20_2021 <- tail(sorted_df[!is.na(sorted_df$Inequality.in.Education..2021.),], 20)

ggplot(bottom_20_2021, aes(x = reorder(Country, Inequality.in.Education..2021.), y = Inequality.in.Education..2021.)) +
  geom_bar(stat = "identity", fill = "salmon") +
  labs(title = "Bottom 20 Countries with Lowest Inequality in Education (2021)") +
  xlab("Country") +
  ylab("Inequality in Education (2021)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# 2020 {data-icon="fa-signal"}

## Row {.tabset .tabset-fade}

### Top 20 countries

```{r}
top_20_2020 <-  head(sorted_df[!is.na(sorted_df$Inequality.in.Education..2020.),], 20)

ggplot(top_20_2020, aes(x = reorder(Country, Inequality.in.Education..2020.), y = Inequality.in.Education..2020.)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Inequality in Education by Country (2020)") +
  xlab("Country") +
  ylab("Inequality in Education (2020)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Bottom 20 countries

```{r}
bottom_20_2020 <- tail(sorted_df[!is.na(sorted_df$Inequality.in.Education..2020.),], 20)

ggplot(bottom_20_2020, aes(x = reorder(Country, Inequality.in.Education..2020.), y = Inequality.in.Education..2020.)) +
  geom_bar(stat = "identity", fill = "salmon") +
  labs(title = "Bottom 20 Countries with Lowest Inequality in Education (2020)") +
  xlab("Country") +
  ylab("Inequality in Education (2020)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# 2010 {data-icon="fa-signal"}

## Row {.tabset .tabset-fade}

### Top 20 countries

```{r}
top_20_2010 <-  head(sorted_df[!is.na(sorted_df$Inequality.in.Education..2010.),], 20)

ggplot(top_20_2010, aes(x = reorder(Country, Inequality.in.Education..2010.), y = Inequality.in.Education..2010.)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Inequality in Education by Country (2010)") +
  xlab("Country") +
  ylab("Inequality in Education (2010)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Bottom 20 countries

```{r}
bottom_20_2010 <- tail(sorted_df[!is.na(sorted_df$Inequality.in.Education..2010.), ], 20)

ggplot(bottom_20_2010, aes(x = reorder(Country, Inequality.in.Education..2010.), y = Inequality.in.Education..2010.)) +
  geom_bar(stat = "identity", fill = "salmon") +
  labs(title = "Bottom 20 Countries with Lowest Inequality in Education (2010)") +
  xlab("Country") +
  ylab("Inequality in Education (2010)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# 2015 {data-icon="fa-signal"}

------------------------------------------------------------------------

## Column {.tabset data-width="400,"}

### Top 20 countries

```{r}
bottom_20_2015 <- head(sorted_df[!is.na(sorted_df$Inequality.in.Education..2015.), ], 20)

ggplot(bottom_20_2015, aes(x = reorder(Country, Inequality.in.Education..2015.), y = Inequality.in.Education..2015.)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Bottom 20 Countries with Lowest Inequality in Education (2015)") +
  xlab("Country") +
  ylab("Inequality in Education (2015)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Bottom 20 countries

```{r}
bottom_20_2015 <- tail(sorted_df[!is.na(sorted_df$Inequality.in.Education..2015.), ], 20)

ggplot(bottom_20_2015, aes(x = reorder(Country, Inequality.in.Education..2015.), y = Inequality.in.Education..2015.)) +
  geom_bar(stat = "identity", fill = "salmon") +
  labs(title = "Bottom 20 Countries with Lowest Inequality in Education (2015)") +
  xlab("Country") +
  ylab("Inequality in Education (2015)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# 3D Scatterplot

## \## Column {.tabset data-width="400,"}

### Human development group

```{r}
inequality_data_3d <- df[, c("Country", "Inequality.in.Education..2010.", "Inequality.in.Education..2015.", "Inequality.in.Education..2021.", "HDI.Rank..2021.")]

plot_3d <- plot_ly(inequality_data_3d, x = ~`Inequality.in.Education..2010.`, y = ~`Inequality.in.Education..2015.`, z = ~`Inequality.in.Education..2021.`, 
                   text = ~Country, color = ~`HDI.Rank..2021.`,
                   type = 'scatter3d', mode = 'markers')

layout_3d <- list(
  scene = list(xaxis = list(title = "Inequality 2010"),
               yaxis = list(title = "Inequality 2015"),
               zaxis = list(title = "Inequality 2021")),
  margin = list(l = 0, r = 0, b = 0, t = 0)
)

plot_3d <- plot_3d %>% layout(layout_3d)
plot_3d

```

### Scatter Plot of ISO3, Country, and UNDP Developing Region

```{r}
fig <- plot_ly(data = df, x = ~ISO3, y = ~Country, z = ~UNDP.Developing.Regions, 
               type = 'scatter3d', mode = 'markers', color = ~UNDP.Developing.Regions)

fig <- fig %>% layout(scene = list(aspectmode = "cube"),
                     title = "3D Scatter Plot of ISO3, Country, and UNDP Developing Region")



fig

```



# Inequality for year

## column {data-width="650"}

```{r}
countries <- c(
  "AFG", "AGO", "ALB", "AND", "ARE", "ARG", "ARM", "ATG", "AUS", "AUT",
  "AZE", "BDI", "BEL", "BEN", "BFA", "BGD", "BGR", "BHR", "BHS", "BIH",
  "BLR", "BLZ", "BOL", "BRA", "BRB", "BRN", "BTN", "BWA", "CAF", "CAN",
  "CHE", "CHL", "CHN", "CIV", "CMR", "COD", "COG", "COL", "COM", "CPV",
  "CRI", "CUB", "CYP", "CZE", "DEU", "DJI", "DMA", "DNK", "DOM", "DZA",
  "ECU", "EGY", "ERI", "ESP", "EST", "ETH", "FIN", "FJI", "FRA", "FSM",
  "GAB", "GBR", "GEO", "GHA", "GIN", "GMB", "GNB", "GNQ", "GRC", "GRD",
  "GTM", "GUY", "HKG", "HND", "HRV", "HTI", "HUN", "IDN", "IND", "IRL",
  "IRN", "IRQ", "ISL", "ISR", "ITA", "JAM", "JOR", "JPN", "KAZ", "KEN",
  "KGZ", "KHM", "KIR", "KNA", "KOR", "KWT", "LAO", "LBN", "LBR", "LBY",
  "LCA", "LIE", "LKA", "LSO", "LTU", "LUX", "LVA", "MAR", "MCO", "MDA",
  "MDG", "MDV", "MEX", "MHL", "MKD", "MLI", "MLT", "MMR", "MNE", "MNG",
  "MOZ", "MRT", "MUS", "MWI", "MYS", "NAM", "NER", "NGA", "NIC", "NLD",
  "NOR", "NPL", "NRU", "NZL", "OMN", "PAK", "PAN", "PER", "PHL", "PLW",
  "PNG", "POL", "PRK", "PRT", "PRY", "PSE", "QAT", "ROU", "RUS", "RWA",
  "SAU", "SDN", "SEN", "SGP", "SLB", "SLE", "SLV", "SMR", "SOM", "SRB",
  "SSD", "STP", "SUR", "SVK", "SVN", "SWE", "SWZ", "SYC", "SYR", "TCD",
  "TGO", "THA", "TJK", "TKM", "TLS", "TON", "TTO", "TUN", "TUR", "TUV",
  "TZA", "UGA", "UKR", "URY", "USA", "UZB", "VCT", "VEN", "VNM", "VUT",
  "WSM", "YEM", "ZAF", "ZMB", "ZWE"
)

selectInput("country", "Choose a country:", choices = countries)
 
 renderPlot({
selected_country <- input$country
selected_year <- input$year

filtered_data <- df[df$ISO3 == selected_country, ]
selected_group <- filtered_data$Human.Development.Groups
bar_colors <- group_colors[selected_group]
bar_data <- as.numeric(unlist(filtered_data[, 6:17]))

years <- 2010:2021
barplot(bar_data, names.arg = years, main = paste("Inequality in Education for", selected_country), col = bar_colors)


  })
    
```
### legend 
```{r}
   group_colors <- c("Very High" = "green", "High" = "lightblue", "Medium" = "orange", "Low" = "red")
 renderPlot({
    legend("top", legend = names(group_colors), fill = group_colors, title = "Development Group")
 })
```
# Inequality

```{r}
df_long <- tidyr::gather(df, key = "Year", value = "Inequality", -ISO3, -Country, -Human.Development.Groups, -UNDP.Developing.Regions, -HDI.Rank..2021.)
df_long$Year <- as.numeric(gsub("Inequality.in.Education..", "", df_long$Year))
p <- ggplot(df_long, aes(x = as.factor(Year), y = Inequality, fill = as.factor(Year))) +
  geom_boxplot() +
  labs(title = "Inequality in Education Over Time",
       y = "Inequality Value",
       x = "Year") +
  theme_minimal()
p <- ggplotly(p)
p

```


# Correlation Heatmap

## Column {data-width="650"}

### Heatmap {data-height="1000"}

```{r}
df_cleaned <- na.omit(df)
inequality_indicators <- df_cleaned[, c("Inequality.in.Education..2010.",
                                "Inequality.in.Education..2011.", "Inequality.in.Education..2012.",
                                "Inequality.in.Education..2013.", "Inequality.in.Education..2014.",
                                "Inequality.in.Education..2015.", "Inequality.in.Education..2016.",
                                "Inequality.in.Education..2017.", "Inequality.in.Education..2018.",
                                "Inequality.in.Education..2019.", "Inequality.in.Education..2020.",
                                "Inequality.in.Education..2021.")]

correlation_matrix <- cor(inequality_indicators)
years <- gsub("Inequality.in.Education..(\\d+)\\.", "\\1", colnames(inequality_indicators))

renderPlotly({
    heatmap_plot <- plot_ly(z = correlation_matrix, type = "heatmap",
                            colorscale = cbind(seq(0, 1, length = 100), c("white", "lightblue", "blue", "darkblue")),
                            x = years,
                            y = years,
                            zauto = FALSE, zmin = -1, zmax = 1) %>%
      layout(title = "Correlation Heatmap",
             xaxis = list(title = "Indicators"),
             yaxis = list(title = "Indicators"))
    
    heatmap_plot
  })
```

### Interpretation

-   This colorscale is used to represent the strength of correlations, with darker shades of blue indicating stronger positive correlations and darker shades of red indicating stronger negative correlations.
